### 2014-05-21 AE_Reponse.tex

\documentclass[11pt,notitlepage,onecolumn]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{makeidx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{chicago}
\usepackage[singlespacing]{setspace}

\setcounter{MaxMatrixCols}{10}

\setlength{\textheight}{9in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\oddsidemargin}{0.0in}

\newcommand{\noi}{\noindent}
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\singlespacing

\baselineskip0.26in

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\pagebreak

\begin{center}
\textbf{\Large Response to the Associate Editor} 
\medskip

{\large Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs}
\medskip

{\footnotesize by D. Love and G. Bayraksan}
\end{center}

\bigskip

\bigskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We thank the Associate Editor for the helpful comments and suggestions. 
We incorporated virtually all of the recommended changes and believe the paper is improved as a result. 
Below, we go through in detail the suggestions of the Associate Editor.
\medskip

\begin{itemize}
\item[]\textit{The paper is not clear in laying out a publishable contribution, and the motivation of the work is not clear.}
\end{itemize}

\noi
We revised the introduction to clearly reflect the contributions of the paper and the motivation for the work, based on your comments and the referees.
\medskip

\begin{itemize}
\item[] \textit{The experiments show that the confidence interval average widths and observed coverage probabilities are the same with overlapping batch means; the only difference is in decreased variability of the confidence interval widths. 
That is, the variance of the variance estimator is reduced. 
This second-order benefit should be weighed against the increased computation cost (not to mention complexity and restricted applicability) of the overlapping batch means method.}
\end{itemize}

\noi  
xxx  Hmm We need to be careful addressing this comment but M1 of ref1 could help.
\medskip

\begin{itemize}
\item[] \textit{The issue of computational efficiency is not addressed  in the paper. 
In this setting the efficiency is often defined as the reciprocal of the computation time multiplied by the variance. 
If this quantity were compared (overlapping batch means versus batch means) in the numerical experiments, I am curious if there are any examples where the efficiency is improved.}
\end{itemize}

\noi 
xxxx Need to discuss
\medskip

\begin{itemize}
\item[] \textit{The authors could argue the publishability of the work by establishing improved efficiency for some examples, or providing some other argument for the usefulness of the results.  
A link between less variable confidence interval widths and improved stochastic optimization efficiency would be helpful.}
\end{itemize}

\noi
xxx  Is this really a new comment?  Looks like just continuation of the prev. comment. 
Also, seems like this is what M1 of ref1 is asking us to do. 
Once we have those results, we can argue more.
\medskip

\bigskip

\noi {\large Minor comments:}
\medskip

\begin{itemize}
\item[] \textit{2.38. The meaning of the set $X$ being independent of the random variable $\xi$ is not clear to me.}
\end{itemize}

\noi 
xxx

\begin{itemize}
\item[] \textit{4.24. \ldots data are generated \ldots}
\end{itemize}

\noi 
xxx

\begin{itemize}
\item[] \textit{19.11 \ldots such a scheme \ldots}
\end{itemize}

\noi
xxx

\end{document}