### 2014-05-21 EinC_Reponse.tex

\documentclass[11pt,notitlepage,onecolumn]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{makeidx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{chicago}
\usepackage[singlespacing]{setspace}

\setcounter{MaxMatrixCols}{10}

\setlength{\textheight}{9in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\oddsidemargin}{0.0in}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\singlespacing

\baselineskip0.26in

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\pagebreak

\begin{center}
\textbf{\Large Response to the Editor-in-Chief} \medskip

{\large Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs} \medskip

{\footnotesize by D. Love and G. Bayraksan}
\end{center}

\bigskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We thank the Editor-in-Chief for the helpful comments and suggestions. 
We incorporated all the recommended changes and believe the paper is improved as a result. 
Below, we go through in detail the suggestions of the Editor-in-Chief.\medskip

\begin{itemize}
\item[] \textit{Your paper has a lot of overlap with the 2011 Winter Simulation paper  (which you also state in your paper). 
This refers to research content as well as text. 
If you revise your paper, the word for word copying should be reduced \ldots}
\end{itemize}

\noindent xxx \ \ 
 We incorporated virtually all the recommended changes suggested by you, the associate editor, and the referees. 
We very much appreciate everyone's time and constructive comments. 
We believe the paper is significantly improved and the wording of the paper, compared to the Winter Simulation Conference paper, is changed as a result of the revision.
We also went through the paper and tried to change the text to reduce word for word copying. 
  \medskip

\begin{itemize}
\item[]\textit{Even more important, in clarifying your contribution (as already asked by the reviewers), you should also shortly explain what new research results are presented in the  TOMACS paper in comparison to the WSC paper}
\end{itemize}

\noindent
xxx Do xxx  We added a paragraph (??) in the introduction, detailing what has been done over \ldots

\end{document}