### 2014-07-11 EinC_Reponse.tex

\documentclass[11pt,notitlepage,onecolumn]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{makeidx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{chicago}
\usepackage[singlespacing]{setspace}

\setcounter{MaxMatrixCols}{10}

\setlength{\textheight}{9in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\oddsidemargin}{0.0in}

\newcommand{\noi}{\noindent}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\singlespacing

\baselineskip0.26in

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\pagebreak

\begin{center}
\textbf{\Large Response to the Editor-in-Chief} \medskip

{\large Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs} \medskip

{\footnotesize by D. Love and G. Bayraksan}
\end{center}

\bigskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noi
We thank the Editor-in-Chief for the helpful comments and suggestions. 
We incorporated all the recommended changes and believe the paper is improved as a result. 
Below, we go through in detail the suggestions of the Editor-in-Chief.\medskip

\begin{itemize}
\item[] \textit{Your paper has a lot of overlap with the 2011 Winter Simulation paper  (which you also state in your paper). 
This refers to research content as well as text. 
If you revise your paper, the word for word copying should be reduced \ldots}
\end{itemize}

\noindent  
We incorporated all the recommended changes suggested by you, the associate editor, and the referees and believe the paper is significantly improved as a result.
We very much appreciate everyone's time and constructive comments. 
As a result of the revision, the wording and content of the paper, compared to the Winter Simulation Conference Proceedings version, has partly changed.
We also edited the paper to improve exposition and tried to cut back word for word copying during this process.  
\medskip

\begin{itemize}
\item[]\textit{Even more important, in clarifying your contribution (as already asked by the reviewers), you should also shortly explain what new research results are presented in the  TOMACS paper in comparison to the WSC paper (please note TOMACS expects about 30\% new research).}
\end{itemize}

\noindent
We added a paragraph in Section 1 (Introduction) detailing what has been done over the WSC proceedings paper. 
Please see the first paragraph on page 5 of the revised manuscript.\bigskip

\noi 
Specifically, we now show (i) almost sure consistency of the variance estimator (earlier this was mean-square consistency), (ii) asymptotic validity of the resulting confidence intervals, and (iii) that overlapping MRP variance estimators have asymptotically lower variances compared to (nonoverlapping) MRP variance estimator. 
These new theoretical results are due to a stronger assumption made in the paper submitted to TOMACS, which was not present in the WSC version. 
We also present a number of new computational results. 
For instance, we (iv) provide computations on two additional test problems, including a large-scale problem, and (v) study computational efficiency of the proposed method (added as a result of the revision).

\end{document}